## Loading and preprocessing the data
```{r loading, echo=TRUE}
library(scales)
library(ggplot2)
library(plyr)
if (!file.exists("data")){
        dir.create("data")
}
activity <- read.csv("./data/activity.csv")
str(activity)
summary(activity)
activity$steps  <-  as.numeric(activity$steps);
activity$interval  <- as.numeric(activity$interval);
activity$date  <- as.Date(strptime(activity$date, format = "%Y-%m-%d"))
```

## What is mean total number of steps taken per day?
```{r total_number, echo=TRUE}
activityDay  <- aggregate(steps ~ date, activity, sum)
ggplot(activityDay, aes(x=steps)) + 
        geom_histogram(binwidth = 1500) +
        labs(title="histogram of the total number of steps taken each day") 
mean(activityDay$steps,na.rm=T)
median(activityDay$steps,na.rm=T)
```

## What is the average daily activity pattern?
```{r daily_activity, echo=TRUE}
fiveMinutesData  <- aggregate(steps ~ interval, activity, mean)
ggplot(fiveMinutesData, aes(x = interval, y = steps, color = "red")) + 
        geom_line() +
        geom_line(aes(y=mean(fiveMinutesData$steps, na.rm=T), color="blue")) +
        scale_colour_manual("legend", values = c("red","blue"), labels = c("average of all days", "average per 5-minutes")) +
        labs(title="Average Daily Activity", x="Interval", y="Number of steps")

interval_max_step <- fiveMinutesData[ fiveMinutesData$steps == max(fiveMinutesData$steps),][[1]]
```
The `r interval_max_step`  5-minute interval, on average across all the days in the dataset contains the maximum number of steps

## Imputing missing values
```{r imputing_missing_values, echo=TRUE}
# Calculate and report the total number of missing values 
sapply(activity, function(x) sum(is.na(x)))
#Devise a strategy for filling in all of the missing values : 
# mean for the same 5-minute interval identifier
activityWithoutNA <- ddply(activity, .(interval), function(df) {
        df$steps[is.na(df$steps)] <- mean(df$steps, na.rm=TRUE); 
        return(df)
        })
activityDay  <- aggregate(steps ~ date, activityWithoutNA, sum)
ggplot(activityDay, aes(x=steps)) + 
        geom_histogram(binwidth = 1500) +
        labs(title="histogram of the total number of steps taken each day") 
mean(activityDay$steps,na.rm=T)
median(activityDay$steps,na.rm=T)
summary(activityWithoutNA)
```
the main is the same while the median differs slightly. I don't observe a visible impact of imputing missing data on the estimates of the total daily number of steps

## Are there differences in activity patterns between weekdays and weekends?
```{r difference_weekdays_weekends, echo=TRUE}
weekday  <- weekdays(activity$date)
weekday[ !weekday %in% c("Samedi", "Dimanche")]  <- "weekday"
weekday[ weekday %in% c("Samedi", "Dimanche")]  <- "weekend"
activity$weekday  <- as.factor(weekday)

activityWeekDay  <- aggregate(steps ~ interval + weekday , activity, mean)
ggplot(activityWeekDay, aes(x = interval, y = steps)) + 
        geom_line(color = "blue") + 
        facet_wrap(~weekday, nrow = 2, ncol = 1) + 
        labs(title="Differences in activity patterns between weekdays and weekends", x = "Interval", y = "Number of steps") + 
        theme_bw()
```
